Combinatorial Proofs and Decomposition Theorems for First-order Logic

### 2021-04-26 LICS2021-final-with-appendices-for-arxiv.tex
final-with-app

\IEEEoverridecommandlockouts
%\IEEEpubid{\makebox[\columnwidth]{\emph{[Long version of the LICS 2021 paper, with proofs in the appendix.]}\hfill}
%    \hspace{\columnsep}\makebox[\columnwidth]{ }}
\IEEEpubid{\makebox[\columnwidth]{978-1-6654-4895-6/21/\$31.00~
\copyright2021 IEEE \hfill} \hspace{\columnsep}\makebox[\columnwidth]{ }}

@@

\newcommand\peircescale{.75}

@@

\newcommand\cpscale{.9}

@@

\\[13ex]

@@

\\[1ex]

@@

\\[10ex]

@@

\\[2ex]

@@

\begin{proof}
  First, we can replace every instance of the $\fff$-rule in $\Deri$ by
  $\weakr$. Then the instances of $\weakr$ and $\conr$ are replaced by
  $\wD$ and $\cD$, which can then be permuted down.
  %Details are in
  %Appendix~\ref{app:LK1-decompose}.
\end{proof}

@@

\begin{proof}
  First, observe that the $\exists$ rule can be permuted under all the
  other rules since $A\ssubst{x}{t}$ has the same structure as $A$ and
  none of the other rules has a premise of the form $\Scons{\exists
    x.A}$. It suffices now to prove that all rules in $\set{\forall,
    \aiD, \tttD}$ can be permuted over the rules in $\set{\sw, \mix,
    \fequ}$, which is straightforward. For the $\fequ$-instances that
  do not involve the quantifiers, the details can be found
  in~\cite{dissvonlutz}, and the $\fequ$-instances concerning the
  quantifiers are admissible in the system without the $\exists$-rule.
\end{proof}

@@

\begin{proof}
  Permute all $\wD$ and $\wfaD$ instances to the bottom of
  the derivation, then permute all $\cD$ and
  $\cfaD$ below $\set{\me,\mfaD,\mexD}$. This involves a tedious but
  routine case analysis. However, unlike most other rule
  permutations in this paper, this has not been done before in the deep
  inference literature. 
%  For this reason, we give the full case
%  analysis in Appendix~\ref{app:cw-decomposition}.
  This
  Lemma is the reason for the presence of the rules $\wfaD$ and
  $\cfaD$, as without them the permutation cases in~\eqref{eq:wfa-cfa}
  could not be resolved.
\end{proof}

@@

\begin{proof}
  By \cite[Proposition~7.5]{str:ral:tableaux19}, there is a derivation
  $\vlder{\set{\acD,\me,\PE\fequ}}{\DDeri}{\PE
    H}{\PE{(G\substof\phi)}}$.  We can lift $\DDeri$ to a first-order
  derivation in $\set{\acD,\cfaD,\me,\mfaD,\mexD,\fequ}$, in a similar
  way as in the previous lemma.
  %The technical details are in
  %Appendix~\ref{app:surj->cm}.
\end{proof}

@@

%% We conjecture cut-elimination for first-order combinatorial proofs can be achieved via a method that is similar to the 
%% \emph{cut-elimination by resolution} method for the sequent calculusthat via the step to first-
%% \lutz{do we want/can say more here?}

%% \todo{mention CERES}

@@

>>>>

\IEEEoverridecommandlockouts
\IEEEpubid{\makebox[\columnwidth]{\emph{[Long version of the LICS 2021 paper, with full proofs in the appendix.]}\hfill}
    \hspace{\columnsep}\makebox[\columnwidth]{ }}
%\IEEEpubid{\makebox[\columnwidth]{978-1-6654-4895-6/21/\$31.00~
%\copyright2021 IEEE \hfill} \hspace{\columnsep}\makebox[\columnwidth]{ }}

@@

\newcommand\peircescale{.78}

@@

\newcommand\cpscale{.95}

@@

\\[15ex]

@@

\\[3ex]

@@

\\[12ex]

@@

\\[4ex]

@@

\begin{proof}
  First, we can replace every instance of the $\fff$-rule in $\Deri$ by
  $\weakr$. Then the instances of $\weakr$ and $\conr$ are replaced by
  $\wD$ and $\cD$, which can then be permuted down.
  Details are in
  Appendix~\ref{app:LK1-decompose}.
\end{proof}

@@

\begin{proof}
  First, observe that the $\exists$ rule can be permuted under all the
  other rules since $A\ssubst{x}{t}$ has the same structure as $A$ and
  none of the other rules has a premise of the form $\Scons{\exists
    x.A}$. It suffices now to prove that all rules in $\set{\forall,
    \aiD, \tttD}$ can be permuted over the rules in $\set{\sw, \mix,
    \fequ}$, which is straightforward. For $\sw$, $\mix$, and the $\fequ$-instances that
  do not involve the quantifiers, the details can be found
  in~\cite{dissvonlutz}. The $\fequ$-instances concerning the
  quantifiers are admissible if the \hbox{$\exists$-rule} is not present.
\end{proof}

@@

\begin{proof}
  Permute all $\wD$ and $\wfaD$ instances to the bottom of
  the derivation, then permute all $\cD$ and
  $\cfaD$ below $\set{\me,\mfaD,\mexD}$. This involves a tedious but
  routine case analysis. However, unlike most other rule
  permutations in this paper, this has not been done before in the deep
  inference literature. 
  For this reason, we give the full case
  analysis in Appendix~\ref{app:cw-decomposition}.
  This
  Lemma is the reason for the presence of the rules $\wfaD$ and
  $\cfaD$, as without them the permutation cases in~\eqref{eq:wfa-cfa}
  could not be resolved.
\end{proof}

@@

\begin{proof}
  By \cite[Proposition~7.5]{str:ral:tableaux19}, there is a derivation
  $\vlder{\set{\acD,\me,\PE\fequ}}{\DDeri}{\PE
    H}{\PE{(G\substof\phi)}}$.  We can lift $\DDeri$ to a first-order
  derivation in $\set{\acD,\cfaD,\me,\mfaD,\mexD,\fequ}$, in a similar
  way as in the previous lemma.
  The technical details are in
  Appendix~\ref{app:surj->cm}.
\end{proof}

@@

@@

%\end{document}
